Improving credibility and efficiency of decision-making \\ within democratic communities via a statistical quorum

### 2018-04-19 statistical-quorum.tex
More discussion.

issues with quorum
unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after voting
Loosing a quorum often during meetings in spring

Plotting time spent per community member on PeerMind gives insights into the
democratization of decision-making. While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours. We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes). Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.

decline in understanding of the process:

We attribute this to the fact that the use of the web application is
a novelty, including using a novel concept like statistical quorum, and requires more effort
in educating members than a traditional decision making.
Moreover, decision making in person allows one to observe others how they are participating and
learn from them, instead of being alone using the application.
We explore this question also through interviews in Section~\ref{sec:interviews}.

This not only means that PeerMind users were more productive in terms of decisions made, they also spent significantly less time per motion than the control group.

\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.\\
Part of the answer here is honestly just that the admin should tune the threshold (and a few of the other hyperparameters we introduce) to their intuition.  Once you set the parameters, the real benefit is that you have consistency, rather than having to always use your intuition as an admin about evaluating credibility.}

>>>>

Based on quantitative data we can observe that members using PeerMind were more productive in
terms of decisions made.
On the other hand, they also rejected more proposals which shows more involvement in decisions
than just rubber stamping proposals.
They also spent significantly less time per proposal than members in spring semester.
While at the same time they made significantly more votes.
Combined with data from the survey where members observed that in the fall their time
is better spend making decisions, we can find support for our hypothesis that directing
attention of members towards contentious issues using a statistical quorum increases the
participation where it is needed (which can lead to a proposal being rejected) and
increases participation overall while keeping the time (or attention) needed low.

Moreover, Figure~\ref{fig:time} shows that participation is diversified and more different
members participate over what we observed through analysis of minutes from the semester prior.
We believe many members who had spent only 30 minutes using PeerMind would not participate
in the in-person council at all (which is confirmed by much lower attendance and fewer votes).
Furthermore, people who need more time to discuss are given the freedom to do so without wasting
time of those who are not interested in the pertinent discussion.
We have observed that councils in the spring had often issues retaining or even achieving a
quorum which we attribute to this mismatch in how people want to spend their time during
decision making and their time availability.

In the case of the in-person decision making it is also unclear how much time to spend on each
dicussion item: how contentious the discussion really is can be seen only after voting.
Using PeerMind members can observe the voting process in real-time and continue discussing
and even changing their votes if they get convinced to change their poisition.
At the same time using feedback provided by credibility value they can also gauge how
much discussion is needed.

% People changed vote online X times after discussion.

The decline in understanding of the process we see in the survey is expected and understandable.
We attribute this to the fact that the use of the web application is
a novelty, including using a novel concept like statistical quorum, and requires more effort
in educating members than a traditional decision making.
Moreover, decision making in person allows one to observe others how they are participating and
learn from them, instead of being alone using the application.

%\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.\\
%Part of the answer here is honestly just that the admin should tune the threshold (and a few of the other hyperparameters we introduce) to their intuition.  Once you set the parameters, the real benefit is that you have consistency, rather than having to always use your intuition as an admin about evaluating credibility.}
